Notes and other text

### 2023-04-28 doc/manuscript/translocation.qmd
Update translocation.qmd

---
bibliography: translocation.bib
csl: pnas.csl
format: docx
editor: visual
---

@@

There are four conditions that will often determine the feasibility of using individuals from populations that are recovering following disease-induced declines to reestablish extirpated populations.
First, populations that survive following disease emergence evolve increased resistance/tolerance.
Second, increased resistance/tolerance contributes to higher individual fitness in the presence of disease and allows positive population growth.
Third, the fitness benefits gained through selection for increased resistance are maintained beyond the natal environment and across generations, allowing for the establishment of reintroduced populations.
And fourth, reintroduced populations show long-term growth and persistence.
To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021; @mendelson2019].

The mountain yellow-legged (MYL) frog, comprised of the sister species *Rana muscosa* and *Rana sierrae* [@vredenburg2007], is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus [*Batrachochytrium dendrobatidis,* @scheele2019].
Once the most common amphibian in the high elevation portion of the California's Sierra Nevada mountains [USA, @grinnell1924], this frog has disappeared from more than 90% of its historical range during the past century [@vredenburg2007].
Due to the severity of its decline and the increasing probability of extinction, both species are now listed as "endangered" under the U.S.
Endangered Species Act.
In the Sierra Nevada, this decline was initiated by the introduction of non-native trout into the extensive historically-fishless region [@bradford1989; @knapp2000], and the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread [@vredenburg2019] caused additional widespread population extirpations [@vredenburg2010; @rachowicz2006].
These Bd-caused declines are fundamentally different from the fish-caused declines because removal of fish populations is feasible and results in the rapid recovery of frogs [@knapp2007; @vredenburg2004].
In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts [@walker2007], and therefore represents a long-term alteration of invaded ecosystems that amphibians will need to overcome to reestablish populations.

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant/tolerant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].

@@

The splined window analysis identified 58 outlier regions for *F~ST~* and 33 outlier regions for $\pi_{diff}$ (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics and 2 of these shared regions also contained outlier SNPs.
A total of 33 annotated genes were found in the 9 shared regions (SI Dataset 2).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a low $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot SI).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended MHC Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

Estimates of frog survival during the 1 year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate SI), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter SI), and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].

@@

Results from frog translocations indicated that adult survival was often relatively high and that most populations showed evidence of successful reproduction and recruitment (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the 2-16 years of site-specific CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).
Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability $\omega$ of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$, recruitment probability $\omega$, and female fecundity *F* showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
Note that we assumed $\sigma_{A_R}$ was equivalent to the survival probability of translocated adults $\sigma_{A_T}$, except for three populations where we had \> 5 years of CMR data and recovering populations that together allowed us to estimate $\sigma_{A_R}$ and $\sigma_{A_T}$ separately (@fig-compare_surv_probs SI).

@@

Disease-induced population declines are impacting global biodiversity [@daszak2000], and strategies to recover affected species are lacking.
Recovery in this context is likely to be particularly challenging because typically recovery will need to occur in the presence of ongoing disease.
One potential strategy, and the one we employ in the current study, utilizes populations of resistant/tolerant individuals as a source of animals with which to reestablish extirpated populations in the presence of disease.
Although previously proposed [@mendelson2019], this strategy has rarely been successfully implemented (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (i) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease (Bd) and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (ii) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing Bd infection.
Although these results are generally consistent with expectations under evolutionary rescue [@carlson2014], obtaining strong evidence in wild animal populations for each of the associated causal links (i.e., selection for adaptive alleles $\rightarrow$ increase in host fitness in the presence of the pathogen $\rightarrow$ positive population growth and recovery) is difficult and perhaps impossible.
Nonetheless, in the following narrative, we use evolutionary rescue as an informative, albeit tentative framework.

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified substantial differences in allele frequencies and larger genomic regions between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the major histocompatibility complex (MHC) plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection contains the gene, GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014].
This gene was also strongly linked to disease in salmon, explaining a remarkable 20% of the resistance phenotype [@robledo2020; @robledo2018].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
In the SI, we provide additional details on these and other genes identified in our study as under selection <!--# Retain or delete? -->.
Collectively, these results suggest that the documented natural recovery of MYL frogs following Bd-caused declines [@knapp2016] is consistent with evolutionary rescue.

@@

In the splined window analysis, we identified outlier regions using *F~ST~* and differences in nucleotide diversity ($\pi_{diff}$) between naive and recovering populations.
First, we calculated per-site *F~ST~* between the naive and recovering individuals for all bi-allelic SNPs in the 30 largest contigs (98% of all SNPs) using VCFtools [@danecek2011].
Next, we calculated per-site $\pi$ separately for individuals from the naive and recovering populations using VCFtools, then calculated $\pi_{diff}$ for each population ($\pi_{diff} = \pi_{naive} - \pi_{recovering}$).
We concatenated the values for *F~ST~* and $\pi_{diff}$ in order of size-sorted chromosome number and adjusted the SNP position based on the relative position in the genome.
We then used the WINGEN R package [@bishop2023] to conduct a splined discrete window analysis for *F~ST~* and $\pi_{diff}$.
This method calculates where non-overlapping window boundaries should occur by identifying inflection points in the fitted spline, therefore balancing false positive and false negative results that occur using other window-based methods [@beissinger2015] .
This method also calculates a W-statistic allowing for outlier identification.
We identified outliers as those with a W-statistic greater than 4 standard deviations above the mean for F~ST~ or above/below the mean for $\pi_{diff}$.
Outlier regions were then identified as those that were outliers in both analyses, meaning that they showed 1) high differentiation between naive and recovering populations, and 2) differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and retrieved annotation for that region using blast.

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ ("$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile \$p\_{J1}\$., and the probability of a juvenile successfully recruiting to an adult $\omega$.

@@

![Results from the splined window analysis showing outlier windows for $\pi_{diff}$ (A) and *F~ST~* (B). Outlier regions for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier regions for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from (A) showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from (B) showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 8 annotated genes known to occur in the extended MHC Class I and II regions.](figures/splinewindow_manhattan.png){#fig-spline-manhattan width="450"}

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (indicated by the dashed line). The red line shows where $\lambda = 1$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript1.jpg){#fig-viability}

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:--------------------------------------------------------------------------|:---------------------|:-------------------------------------------------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |

>>>>

---
bibliography: translocation.bib
csl: pnas.csl
format: docx
editor: visual
editor_options:
  markdown:
    wrap: sentence
---

@@

There are four conditions that will often determine the feasibility of using individuals from populations that are recovering following disease-induced declines to reestablish extirpated populations.
First, populations that survive following disease emergence evolve increased resistance/tolerance.
Second, increased resistance/tolerance contributes to higher individual fitness in the presence of disease and allows positive population growth.
Third, the fitness benefits gained through selection are maintained beyond the natal environment and across generations, allowing for the establishment of reintroduced populations.
And fourth, reintroduced populations show long-term growth and persistence.
To date, few of these conditions are documented, and reintroducing hosts from recovering populations is largely untested [@scheele2021; @mendelson2019].

The mountain yellow-legged (MYL) frog, composed of the sister species *Rana muscosa* and *Rana sierrae* [@vredenburg2007], is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus [*Batrachochytrium dendrobatidis,* @scheele2019].
Once the most common amphibian in the high elevation portion of the California's Sierra Nevada mountains [USA, @grinnell1924], this frog has disappeared from more than 90% of its historical range during the past century [@vredenburg2007].
Due to the severity of its decline and the increasing probability of extinction, both species are now listed as "endangered" under the U.S.
Endangered Species Act.
In the Sierra Nevada, this decline was initiated by the introduction of non-native trout into the extensive historically-fishless region [@bradford1989; @knapp2000], and the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread [@vredenburg2019] caused additional widespread population extirpations [@vredenburg2010; @rachowicz2006].
These Bd-caused declines are fundamentally different from the fish-caused declines because removal of fish populations is feasible and results in the rapid recovery of frogs [@knapp2007; @vredenburg2004].
In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts [@walker2007], and therefore represents a long-term alteration of invaded ecosystems that amphibians will need to overcome to reestablish populations.

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are extirpated following Bd arrival [@vredenburg2010], some populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant/tolerant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].

@@

The splined window analysis identified 58 outlier regions for *F~ST~* and 33 outlier regions for $\pi_{diff}$ (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics and 2 of these shared regions also contained outlier SNPs.
A total of 33 annotated genes were found in the 9 shared regions (SI Dataset 2).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a low $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot SI).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended major histocompatibility complex (MHC) Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

Estimates of frog survival during the 1 year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate SI), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation (@fig-bdload-beforeafter SI), and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].

@@

Results from frog translocations indicated that adult survival was often relatively high and that most populations showed evidence of successful reproduction and recruitment (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the 2-16 years of site-specific CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).
Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability $\omega$ of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$, recruitment probability $\omega$, and female fecundity *F* showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
Note that we assumed $\sigma_{A_R}$ was equivalent to the survival probability of translocated adults $\sigma_{A_T}$, except for three populations where we had \> 5 years of CMR data and recovering populations that together allowed us to estimate $\sigma_{A_R}$ and $\sigma_{A_T}$ separately (@fig-compare_surv_probs SI).
<!--# Consider condensing the above paragraph and providing details in SI. Next paragraph topic sentence could also provide a more direct statement of key results, e.g., something like "Even when incorporating demographic and environmental stochasticity, populations with high adult survival are likely to persist over a 50 year time horizon." -->

@@

Disease-induced population declines are impacting global biodiversity [@daszak2000], and strategies to recover affected species are lacking.
Here, we used populations of resistant/tolerant individuals to reestablish extirpated populations, despite ongoing disease.
Although previously proposed [@mendelson2019], this strategy has rarely been successfully implemented (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (i) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease (Bd) and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (ii) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing Bd infection.
Although these results are generally consistent with expectations under evolutionary rescue [@carlson2014], obtaining strong evidence in wild animal populations for each of the associated causal links (i.e., selection for adaptive alleles $\rightarrow$ increase in host fitness in the presence of the pathogen $\rightarrow$ positive population growth and recovery) is difficult and perhaps impossible.
Nonetheless, in the following narrative, we use evolutionary rescue as an informative, albeit tentative framework.
<!-- I am wondering whether it is worth digging a bit deeper into the "difficult and perhaps impossible" task alluded to above. e.g., what would it take to increase confidence that evolutionary rescue is at play, what alternative explanations are still credible given the available data -->

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified substantial differences in allele frequencies and larger genomic regions between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection contains the gene, GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014].
This gene was also strongly linked to disease in salmon, explaining a remarkable 20% of the resistance phenotype [@robledo2020; @robledo2018].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
In the SI, we provide additional details on these and other genes identified in our study as under selection <!--# Retain or delete? -->.

@@

In the splined window analysis, we identified outlier regions using *F~ST~* and differences in nucleotide diversity ($\pi_{diff}$) between naive and recovering populations.
First, we calculated per-site *F~ST~* between the naive and recovering individuals for all bi-allelic SNPs in the 30 largest contigs (98% of all SNPs) using VCFtools [@danecek2011].
Next, we calculated per-site nucleotide diversity $\pi$ separately for individuals from the naive and recovering populations using VCFtools, then calculated $\pi_{diff}$ for each population ($\pi_{diff} = \pi_{naive} - \pi_{recovering}$).
We concatenated the values for *F~ST~* and $\pi_{diff}$ in order of size-sorted chromosome number and adjusted the SNP position based on the relative position in the genome.
We then used the WINGEN R package [@bishop2023] to conduct a splined discrete window analysis for *F~ST~* and $\pi_{diff}$.
This method calculates where non-overlapping window boundaries should occur by identifying inflection points in the fitted spline, therefore balancing false positive and false negative results that occur using other window-based methods [@beissinger2015] .
This method also calculates a W-statistic allowing for outlier identification.
We identified outliers as those with a W-statistic greater than 4 standard deviations above the mean for F~ST~ or above/below the mean for $\pi_{diff}$.
Outlier regions were then identified as those that were outliers in both analyses, meaning that they showed 1) high differentiation between naive and recovering populations, and 2) differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and retrieved annotation for that region using blast.

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ (the subscript "$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile $p_{J1}$., and the probability of a juvenile successfully recruiting to an adult $\omega$.

@@

![Results from the splined window analysis showing outlier windows for the difference in nucleotide diversity $\pi_{diff}$ (A) and *F~ST~* (B). Outlier regions for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier regions for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from (A) showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from (B) showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 8 annotated genes known to occur in the extended MHC Class I and II regions.](figures/splinewindow_manhattan.png){#fig-spline-manhattan width="450"}

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate \$\\lambda\$ for different values of yearly adult survival probability \$\\sigma\_{A_R}\$ and probability of recruitment \$\\omega\$, given the parameterized, deterministic model. Colored points show the predicted \$\\lambda\$ values for the twelve translocated populations when probability of recruitment into the adult class is \$\\omega = 0.3\$ (indicated by the dashed line). The red line shows where \$\\lambda = 1\$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in \$\\omega\$, and different mean values of \$\\omega\$. The dashed line shows where \$\\omega = 0.3\$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript1.jpg){#fig-viability} <!-- Consider a diverging color ramp for panel A, e.g., where lambda=1 is white, lambda>1 is blue, lambda<1 is red -->

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:----------------------------|:-----------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |
